Analysis of the 2005--2015 Earthquake Sequence in Northern Iran Using the Visibility Graph Method

### 2016-08-21 results.tex
improving figure 8

We generate the visibility graph for each sub-regions of interest using the time-magnitude sequences shown in Fig.~\ref{fig:mag-time}. The number of nodes in each graph is the same as the (declustered) number of events shown in Table \ref{tab:seismicity}, and the links between intervisible events are established following the condition in Eq.~\ref{eq:vg}, as in the example shown in Fig.~\ref{fig:vg}. (We do not include a visualization of the complete sequence graphs because the links are so many, it is only practical to visualize the graphs of short sequences). We then collect information about the number of inter-visibility links associated to each event (connectivity degree, $k$) and categorize the events in magnitude bins of size $\Delta M = 0.1$ as mentioned in the Methodology section. Figure \ref{fig:km} shows the scattered distribution of events in the magnitude-connectivity degree plane. The figure shows that, in general, the value of $k$ increases with $M_w$. Figure \ref{fig:km} also shows the result of obtaining linear $k$-$M$ regressions for the datasets and the values of the $k$-$M$ slope for each of the three zones.

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-06} 
	\caption{Scattered distribution of events in the magnitude-connectivity degree plane and the linear regressions obtained for the $k$-$M$ relationship for the three seismic regions in northern Iran. The values next to each of the regression lines corresponds to the slope of each line, which is referred here as the $k$-$M$ slope. The color version of this figure is available only in the electronic edition.}
	\label{fig:km}
\end{figure*}

Next, we examine the relationship between the $b$-values and the from Table \ref{tab:seismicity} and the $k$-$M$ slope values shown in Fig.~\ref{fig:km}. Figure \ref{fig:regression} shows the scattered results for $k-M$ slope versus $b$-value for the three tectonic seismic regions in northern Iran along with the data-points obtained for analysis of the magnitude-time sequences of the Mexican subduction zone \citep{Telesca2013} and the Pannoninan seismic zone \citep{Telesca2014}, as well as other experimental results \citep{Telesca2014-pone}. Figure \ref{fig:regression} also include the results of different linear regressions between $b$ and the $k$-$M$ slope. Each regression shows the addition of new data-points from different studies over time, with the present study adding the latest set of data. Note that the regression improves as new data-points are added over time, which is indicated by the correlation coefficient $R$, whose values are also shown in the figure. This was previously pointed out by \citet{Telesca2014}.

@@

\begin{center}
* \hspace{2em} * \hspace{2em} *
\end{center}

Although the number of events vary significantly between the three zones (see Table \ref{tab:seismicity}, \citet{Telesca2013} observed that the value of the $k$-$M$ slope is not particularly sensitive to the sample size in the sequence---at least not when considering sufficiently large sequences. On the other hand, as we will see below, if the sequence window is sufficiently short, then the $k$-$M$ slope value shows a relative dependence on time and thus provides insight about the variation of the seismicity as the sequence progresses. Note also that the threshold value of $M_c$ is significantly smaller for Azerbaijan than for Alborz or Kopeh-Dagh. This is due in part to the fact that the latter two zones were more seismically active in the time period under consideration. However, according to \citet{Telesca2012}, the threshold magnitude has a minor effect in the graph parameters.

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude to be considered, we randomly picked a significant number of sub-sequences from within the catalog of each region and repeated the analysis. This was done starting with the catalogs before declustering. In total, for each region, we extracted 200 new sub-sequences. The number of events in each sub-sequence was varied randomly but chosen to be large enough to represent the seismic characteristics of each seismic region. In particular, the minimum size of each subsequence was set to be $\geq 150$, whereas the maximum size in the sequence was set to be as large as the original (pre-declustering) catalog. We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), then picked that number of events following the initial event. Next, we determined the value of $M_c$ and $b$ and created the graph to extract the connectivity degree of the events in each sub-sequence. Figure \ref{fig:random}a shows...

@@

% In order to analyze the sensitivity of the catalogs to the number of events and the threshold magnitude, we randomly picked 200 sequences from the catalogs with various sequence sizes. Number of events at each sequence should be low enough to be able to choose 200 random sequence from the catalog (note that the number of events in each catalog, before removing the events with magnitude less than the completeness magnitude, is 271, 1262, and 399 for Azerbaijan, Alborz, and Kopeh Dagh regions, respectively). Also it should be long enough to preserve the seismic characteristic of each tectonic seismic region. Based on mentioned criteria, we choose the minimum size of random windows as 150 events. The maximum size is the size of the entire catalog. The random sequences are consecutive, in other words, we randomly determine the initial event and sequence size, then we pick that size of event starting from the initial event. For each randomly picked magnitude time series,  we estimate $Mc$; then the $k-M$  and $b-value$ are calculated. Fig.~\ref{fig:random} shows the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.

>>>>

We generate the visibility graph for each sub-region of interest using the time-magnitude sequences shown in Fig.~\ref{fig:mag-time}. The number of nodes in each graph is the same as the (declustered) number of events shown in Table \ref{tab:seismicity}, and the links between inter-visible events are established following the condition in equation (\ref{eq:vg}), as in the example shown in Fig.~\ref{fig:vg}. (We do not include a visualization of the complete sequence graphs because the links are so many, it is only practical to visualize the graphs of short sequences). We then collect information about the number of inter-visibility links associated with each event (connectivity degree, $k$) and categorize the events in magnitude bins of size $\Delta M = 0.1$, as mentioned in the Methodology section. Figure \ref{fig:km} shows the scattered distribution of events in the magnitude-connectivity degree plane for each sub-region. The figure shows that, in general, the value of $k$ increases with $M_w$. Figure \ref{fig:km} also shows the results of obtaining linear $k$-$M$ regressions for each dataset and the values of the $k$-$M$ slopes for the three seismic zones.

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-06} 
	\caption{Scattered distribution of events in the magnitude-connectivity degree plane and linear regressions obtained for the $k$-$M$ relationships for the three seismic regions in northern Iran. The value next to each of regression line corresponds to the slope of the line, which is referred here as the $k$-$M$ slope. The color version of this figure is available only in the electronic edition.}
	\label{fig:km}
\end{figure*}

Next, we examine the relationship between the $b$-values from Table \ref{tab:seismicity} and the $k$-$M$ slope values shown in Fig.~\ref{fig:km}. Figure \ref{fig:regression} shows the scattered results for $k-M$ slope versus $b$-value for the three tectonic seismic regions in northern Iran along with the data-points obtained for analysis of the magnitude-time sequences of the Mexican subduction zone \citep{Telesca2013} and the Pannoninan seismic zone \citep{Telesca2014}, as well as other experimental results \citep{Telesca2014-pone}. Figure \ref{fig:regression} also include the results of different linear regressions between the $b$-value and the $k$-$M$ slope. Each regression shows the addition of new data-points from different studies over time, with the present study adding the latest set of data. Note that the regression improves as new data-points are added, which is indicated by the correlation coefficient $R$. This was previously pointed out by \citet{Telesca2014}. The values of the correlation coefficients are also shown in the figure. According to these results, the universal relationship between the $b$-value and the $k$-$M$ slope is given by:
% 
\begin{equation}
	b = 000 + 000 m \, 
	\label{eq:universal.bm}
\end{equation}
% 
where $m$ is the value of the $k$-$M$ slope.

@@

The level of correlation indicated by the values of $R$ in Fig.~\ref{fig:regression} being close to 1 is indicative of the stability of the relationship, which could be considered universal provided that future tests continue to reinforce these results. Another aspect of interest is the stability of the data points themselves. Note that as presented in Fig.~\ref{fig:regression}, the analysis of the sequences shown in Fig.~\ref{fig:mag-time} only contribute one data point per sub-region of interest. Furthermore, each data-point comes from sequences that vary significantly in terms of the number of events and seismic parameters.

\citet{Telesca2013} observed that the value of the $k$-$M$ slope is not particularly sensitive to the sample size in the sequence---at least not when considering sufficiently large sequences. On the other hand, as we will see below, if the sequence window is sufficiently short, then the $k$-$M$ slope value shows a relative dependence on time and thus provides insight about the variation of the seismicity as the sequence progresses. Note also that the threshold value of $M_c$ is significantly smaller for Azerbaijan than for Alborz or Kopeh-Dagh. This is due in part to the fact that the latter two zones were more seismically active in the time period under consideration. However, according to \citet{Telesca2012}, the threshold magnitude has a minor effect in the graph parameters.

To further explore the sensitivity of the graph properties to the number of events in each catalog and the value of the minimum magnitude to be considered, we randomly picked a significant number of sub-sequences from within the initial catalog compiled for each region, and repeated the analysis for each sub-sequence. In total, for each region, we extracted 200 new sub-sequences from the initial (pre-declustering) catalog. The number of events in each sub-sequence was varied randomly but chosen to be large enough to represent the seismic characteristics of each region. In particular, the minimum size of each sub-sequence was set to be $n \geq 150$, whereas the maximum size in the sequence was set to be as large as the original (pre-declustering) catalog. We forced the random sub-sequences to progress positively in time without altering the natural occurrence of events. In other words, we randomly determined the initial event and the sub-sequence size (number of events to be considered), then picked that number of events following the initial earthquake in the sub-sequence. Next, we determined the value of $M_c$ and $b$, created the graph for all events with $M \geq M_c$, and extracted the connectivity degree of the events in each sub-sequence.

Figure \ref{fig:random}a shows scattered points for all the individual $k$-$M$ slope and $b$-value pairs for all the sub-sequences that were randomly generated for each region in northern Iran. Next to it, Figs.~\ref{fig:random}b and \ref{fig:random}c show the

The figure also shows the linear correlation for the

the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.

@@

% Fig.~\ref{fig:random} shows the relationship between  $k-M$  slope and  $b-value$  of the randomly selected data and the statistical parameters for variables. Although changing the number of events and the threshold magnitude is slightly changing the results, however,  the results are fairly well clustered for each tectonic seismic region.
